---
title: "Over Crowding (>1)"
subtitle: "Data source(s): PUMS 2020, 5y (CSV)"
author: "Meg Grzybowski"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output:
  html_document:
    keep_md: yes
    df_print: paged
    toc: yes
    toc_depth: 6
    toc_float: yes
  pdf_document:
    toc: yes
    toc_depth: '6'
---

```{r rmarkdown setup, include=FALSE}
knitr::opts_chunk$set(echo=FALSE, 
                      warning=FALSE, 
                      message=FALSE) # formatting
```

```{r library setup, include=FALSE}
#devtools::install_github("psrc/psrcplot",
 #                         force=TRUE)
library(tidyverse)
library(tidycensus)
library(psrccensus)
library(psrcplot)
library(psrctrends)
library(rlang) #required for psrccensus
library(emmeans) #required for rlang
library(magrittr)
#library(kableExtra)
#library(ggplot2)

library(summarytools) #freq
library(vtable) #summary stats
library(table1)  #nice descriptive summary table
library(scales) #number formatting
library(ggpubr) #graphing - ggarrange fx
library(forcats) #for factor re-leveling
library(plotly) #for interactive charts

#remove.packages("Rcpp")
#install.packages("Rcpp")
#library(Rcpp)
library(odbc) #connect to ElmerGeo
library(DBI) #connect to ElmerGeo
library(sf)
#library(leaflet)
#library(leafem) #home button, part of leaflet/Rcpp
#library(htmlwidgets) #save visuals as html
#library(raster)
#library(ggspatial)
library(lubridate) #year formatting
library(stringr) #add leading zero's
library(reshape2) #formatting data
#library(readxl)
library(RColorBrewer)
library(gridExtra)

install_psrc_fonts()
library(showtext) #trying to fix PSRC font issues
library(sysfonts) #required for showtext
library(showtextdb) #required for showtext

library(dplyr)
#install.packages("srvyr")
library(srvyr)
```

# Introduction
The data set referenced in this script is generated from the American Community Survey (ACS) Public Use Microdata Sample (PUMS) and the Washington Office of Superintendent of Public Instruction [(OSPI)](https://www.k12.wa.us/data-reporting/data-portal). These data sets provide data at the **person-level**, with the ability to look at indicators by the six equity demographic groups of interest. 

# Access data
## PUMS and OSPI data (Elmer)
This data set was compiled from PUMS data. 
```{r equity quintiles in Elmer}
# connect to Elmer
elmer_connection <- dbConnect(odbc::odbc(),
  driver = "SQL Server",
  server = "AWS-PROD-SQL\\Sockeye",
  database = "Elmer",
  trusted_connection = "yes"
  )

# PUMS/OSPI data set
pums_ospi_elmer <- dbReadTable(elmer_connection, SQL("equity.v_tracker_indicators"))

#dbDisconnect(elmer_connection)
```
Looking at the fields in the data set.
```{r}
# explore the data
head(pums_ospi_elmer)
# view the equity demographic groups
unique(pums_ospi_elmer$focus_type)
# view the different indicators available within the data set
unique(pums_ospi_elmer$indicator_type)
# view the different data years available within the data set
unique(pums_ospi_elmer$data_year)
```

# Housing Cost Burden
```{r indicator specific information, include=FALSE}
# set up a variable for analysis
indicator_measure <- 'crowding'
# set up a variable for titles
indicator_title <- 'Overcrowding (>1.5 person/bedroom)'
```

## 1. Explore data
*In this section we make sure that the data set makes sense.*  
```{r}
# clean data set
data_indicator <- pums_ospi_elmer %>% 
  mutate(indicator_attribute = case_when(indicator_attribute == "One person per bedroom or less" |
                                          indicator_attribute == "Between 1 and 1.5 person(s) per bedroom" ~ "Less than 1.5 person(s) per bedroom",
                                          TRUE~ indicator_attribute))%>% 
  dplyr::select(-indicator_fact_id)%>%
  group_by(across(!c(fact_value, margin_of_error)))%>%
  summarise(fact_value = sum(fact_value, na.rm = TRUE),
            margin_of_error = moe_sum(margin_of_error, estimate = fact_value, na.rm = TRUE))

data_full <- data_indicator %>% 
  dplyr::filter(indicator_type ==indicator_measure) %>% 
  dplyr::filter(focus_type !="Total") %>% 
  filter(focus_attribute !="Total") %>%
  filter(indicator_attribute !="Total") %>%
  # distinct(indicator_fact_id, .keep_all = TRUE) %>%
  dplyr::mutate(data_year_yr = format(data_year,format="%Y"))

```

### Data fields
#### *consistent base data*
These fields should be the same across all indicators:

* There should be 5 geographies - the 4 counties and the Region. 
* There should be 6 equity focus categories - POC, income, disability, youth, older adult, and LEP
    + 2 sub-categories per focus group (e.g. people of color, non-people of color)
```{r}
# geographies
num_county <- length(unique(data_full$county)) #5
unique(data_full$county)

# equity focus groups
num_group <- length(unique(data_full$focus_type)) #6
unique(data_full$focus_type)
```
\
\

#### *indicator-specific data*
These fields will vary by indicator:

* Type of metric - this will determine how the data are visualized *(est ="percent" or "currency" or "number" or "share")*
* Number of years (5-year span) - this can vary depending on data availability
* Number of indicator-specific categories - this can vary depending on the indicator of interest, ranging from N/A (*median income*) to multiple levels (*crowding, housing cost burden*)
```{r}
# metric
unique(data_full$fact_type) #share

# years
num_yr <- length(unique(data_full$data_year)) #3
unique(data_full$data_year)
time_period <- '2011, 2016, 2021'

# indicator attributes
num_indatt <- length(unique(data_full$indicator_attribute))
unique(data_full$indicator_attribute)

# calculate expected no. rows
num_row <- num_yr*num_county*num_group*2*num_indatt #540
```

There are **`r num_county`** geographies and **`r num_group`** equity focus groups (each with **2** subgroups). There are **`r num_yr`** years in the data set and the indicator specific field has **`r num_indatt`** attribute(s), which means there should be a total of **`r num_row`** rows.
```{r}
# count number of rows
nrow(data_full) #340
```
<span style="color: #00A7A0">There are some missing data.</span> 
\
\

#### *checking for missing data*
##### Year / geography
```{r}
num_yr_geo <- num_group*2*num_indatt #24
```
If we look at the data by year and geography, there should be **`r num_yr_geo`** entries per year/geography.
```{r}
table(data_full$data_year,
      data_full$county)
```
<span style="color: #00A7A0">
2011 is missing some data for all geographies.
</span> 
\
\

##### Year / equity focus group
```{r}
num_yr_grp <- num_county*2*num_indatt #20
```
If we look at the data by year and focus group, there should be **`r num_yr_grp`** entries per year/focus group.
```{r}
table(data_full$data_year,
      data_full$focus_type)
```
<span style="color: #00A7A0">
The disability category is missing all data in 2011.
</span> 
\
\

##### Year / equity focus sub-group
```{r}
num_yr_subgrp <- num_county*num_indatt #10
```
If we look at the data by year and focus sub-group, there should be **`r num_yr_subgrp`** entries per year/focus sub-group.
```{r}
table(data_full$data_year,
      data_full$focus_attribute)
```
<span style="color: #00A7A0">
The disability category is missing all data in 2011.
</span> 
\
\

##### Year / indicator attribute
```{r}
num_yr_ind <- num_county*num_group*2 #60
```
If we look at the data by year and indicator attribute, there should be **`r num_yr_ind`** entries per year/indicator attribute.
```{r}
table(data_full$data_year,
      data_full$indicator_attribute)
```
<span style="color: #00A7A0">
All data in 2011 is missing 10 entries.
</span> 
\
\

### Numeric data
To check for *0*s and *NULL*s
```{r}
# checking for NULL
summary(data_full$fact_value)
```
<span style="color: #00A7A0">
There are no 0s or nulls.
</span> 
\
\

To look at distribution of all data - not the most useful visual, but provides a sense of the range of values at a high level in one plot. 
\
```{r}
qplot(data_year_yr, fact_value, 
      colour = county, shape = focus_type, 
      data = data_full, 
      main = indicator_title) + theme(axis.title.x = element_blank(),
                                               axis.title.y = element_blank())
```
\
This table includes a lot of information about the data set and helps to show the different levels of each field. It provides another way to check if data are available for all counties and all years, or where there may be gaps in the data set. 
```{r}
data_refined <- data_full %>% 
  dplyr::select(data_year_yr, county, vulnerability, focus_type, focus_attribute,
                indicator_type, indicator_attribute, fact_value)

sumtable(data_refined,
         add.median=TRUE,
         group="county",
         out="return")
```

```{r}
# renaming labels ----
data_clean <- data_full %>% 
  mutate(focus_type_edit = case_when(focus_type=="POC_cat"~"race/ethnicity",
                                     focus_type=="Disability_cat"~"disability status",
                                     focus_type=="LEP_cat"~"English proficiency",
                                     focus_type=="Income_cat"~"income",
                                     focus_type=="Youth_cat"~"youth <18",
                                     focus_type=="Older_cat"~"older adults 65+"))

# re-ordering variables ----
data_clean$focus_attribute_ord <- factor(data_clean$focus_attribute,
                                         levels = c("POC",
                                                    "Non-POC",
                                                    "Low Income",
                                                    "Non-Low Income",
                                                    "With disability",
                                                    "Without disability",
                                                    "Limited English proficiency",
                                                    "English proficient",
                                                    "Household with youth",
                                                    "Household without youth", 
                                                    "Household with older adult",
                                                    "Household without older adult",
                                                    "Total"))

data_clean$county_ord <- factor(data_clean$county,
                                levels = c("King",
                                           "Kitsap",
                                           "Pierce",
                                           "Snohomish",
                                           "Region"))

data_clean$focus_type_ord <- factor(data_clean$focus_type_edit,
                                    levels = c("race/ethnicity",
                                               "income",
                                               "disability status",
                                               "English proficiency",
                                               "youth <18",
                                               "older adults 65+"))
```

```{r}
# indicator specific re-ordering
data_clean$indicator_attribute <- factor(data_clean$indicator_attribute,
                                    levels = c("Less than 1.5 person(s) per bedroom",
                                               "More than 1.5 persons per bedroom"))
```
\

### Data labels, shares
These charts were generated to ensure the labels across years are consistent/make sense. There had been an issue with misassigned labels because tidycensus::pums_variables, i.e. the only digital data dictionary available to associate labels with codes, exists only from 2017 forward. Most variables have had consistent codes, but in cases where the codes have shifted over time, using the 2017 lookup winds up mischaracterizing categories.
\
\
These charts also help to confirm that the shares add up to 100% - only relevant when *indicator_attribute* has more than one category. The *indicator_attribute* for educational attainment includes 2 categories: "Less than a Bachelor's degree" and "Bachelor's degree or higher" 
\
\
*The colors of the charts may not be consistent between the years depending on missing data.* 
\
```{r}
just_region_11 <- data_clean %>% 
  filter(county=="Region") %>% 
  filter(data_year==2011) %>% 
  arrange(focus_attribute_ord, indicator_attribute)
just_region_16 <- data_clean %>% 
  filter(county=="Region") %>% 
  filter(data_year==2016) %>% 
  arrange(focus_attribute_ord, indicator_attribute)
just_region_21 <- data_clean %>% 
  filter(county=="Region") %>% 
  filter(data_year==2021) %>% 
  arrange(focus_attribute_ord, indicator_attribute)

bar_chart <- static_bar_chart(t=just_region_11,
                               x="fact_value", y="focus_attribute_ord",
                               fill="indicator_attribute",
                               est = "percent",
                               pos= "stack",
                               color = "psrc_light", 
                               title = paste0(indicator_title, ": Region (2011)"),
                               subtitle = "by equity focus group",
                               source = "U.S. Census Bureau, American Community Survey (ACS) 5-Year Public Use Microdata Sample (PUMS)")
bar_chart

bar_chart <- static_bar_chart(t=just_region_16,
                               x="fact_value", y="focus_attribute_ord",
                               fill="indicator_attribute",
                               est = "percent",
                               pos= "stack",
                               color = "psrc_light",
                               title = paste0(indicator_title, ": Region (2016)"),
                               subtitle = "by equity focus group",
                               source = "U.S. Census Bureau, American Community Survey (ACS) 5-Year Public Use Microdata Sample (PUMS)")
bar_chart

bar_chart <- static_bar_chart(t=just_region_21,
                               x="fact_value", y="focus_attribute_ord",
                               fill="indicator_attribute",
                               est = "percent",
                               pos= "stack",
                               color = "psrc_light",
                               title = paste0(indicator_title, ": Region (2021)"),
                               subtitle = "by equity focus group",
                               source = "U.S. Census Bureau, American Community Survey (ACS) 5-Year Public Use Microdata Sample (PUMS)")

bar_chart
```

## 2. Visually explore data
### 2a. Scatter plots
*In this section we start to explore the data visually - distribution by the different dimensions within the data set. These plots are helpful to check for outliers and get a higher level understanding of the data in one visual, before slicing the data by geography and equity focus group in the following sections.* 
\
\
The following code will need to be adjusted to fit the fields specific to the data indicator. For housing cost burden, we focus on those with more than 50 percent cost burden. The following code establishes the data frame that the rest of the analysis uses. **If there are fewer than 2 indicator attributes**, this section can be skipped/commented out, but the code will need to be adjusted throughout to call the correct data frame. 
\
\

#### By *indicator_attribute*
```{r}
# separating indicator attributes by levels ----
data_clean_more <- data_clean %>% 
  filter(indicator_attribute=="More than 1.5 persons per bedroom")
data_clean_less <- data_clean %>% 
  filter(indicator_attribute=="Less than 1.5 person(s) per bedroom")

```

```{r}
# plot data
qplot(data_year_yr, fact_value, 
      colour = county, shape = focus_type_ord, 
      data = data_clean_more, 
      main = paste0(indicator_title, ": Region, more than 1.5 persons per bedroom")) + theme(axis.title.x = element_blank(),
                                                          axis.title.y = element_blank()) + ylim(0, 1)

qplot(data_year_yr, fact_value, 
      colour = county, shape = focus_type_ord, 
      data = data_clean_less, 
      main = paste0(indicator_title, ": Region, less than 1.5 persons per bedroom")) + theme(axis.title.x = element_blank(),
                                                          axis.title.y = element_blank()) + ylim(0, 1)
```

#### By Year
```{r}
# separating by year - focusing on only more than 50 percent cost burdened
data_clean11 <- data_clean_more %>% 
  filter(data_year==2011)
data_clean16 <- data_clean_more %>% 
  filter(data_year==2016)
data_clean21 <- data_clean_more %>% 
  filter(data_year==2021)
```

```{r}
# plot data, adjust y limits based on data set
qplot(focus_type_ord, fact_value, 
      colour = county, shape = vulnerability, 
      data = data_clean11, 
      main = paste0(indicator_title, ": 2011")) + theme(axis.title.x = element_blank(),
                                                        axis.title.y = element_blank()) + ylim(0, 0.65)

qplot(focus_type_ord, fact_value, 
      colour = county, shape = vulnerability, 
      data = data_clean16,
      main = paste0(indicator_title, ": 2016")) + theme(axis.title.x = element_blank(),
                                                        axis.title.y = element_blank()) + ylim(0, 0.65)

qplot(focus_type_ord, fact_value, 
      colour = county, shape = vulnerability, 
      data = data_clean21, 
      main = paste0(indicator_title, ": 2021")) + theme(axis.title.x = element_blank(),
                                                        axis.title.y = element_blank()) + ylim(0, 0.65)
```
\

### 2b. Facets by geography
*In this section we explore trends by different groups with MOEs. These charts help to show any missing data by geography, year, or focus group/subgroup.*
```{r, include=FALSE}
facet_by_geography <- function(geography, focus.text){
  # filter data
  data_clean_geo <- data_clean_more %>% 
    filter(county==geography)%>%
    filter(data_year_yr==2021)
  
  # make chart, adjust estimate type, subtitle
  static_facet_column_chart(t=data_clean_geo,
                            x="data_year_yr", y="fact_value",
                            fill="vulnerability", facet="focus_type_ord",
                            color="psrc_light",
                            est ="percent",
                            moe = "margin_of_error",
                            dec=2,
                            scales="fixed",
                            ncol=3,
                            title=paste0(indicator_title, focus.text),
                            subtitle="More than 1.5 persons per bedroom",
                            source="U.S. Census Bureau, American Community Survey (ACS) 5-Year Public Use Microdata Sample (PUMS)")
}
```

```{r}
facet_by_geography("Region", ": Region")
facet_by_geography("King", ": King County")
facet_by_geography("Kitsap", ": Kitsap County")
facet_by_geography("Pierce", ": Pierce County")
facet_by_geography("Snohomish", ": Snohomish County")
```
\
\

## 3. Developing visuals
*In this section we further develop the draft visuals for communicating the results and supporting the narrative for the Equity Tracker webpages. These charts are slightly more refined by slicing the data by geography and equity focus group. The line charts don't include MOEs, but they help make connections between the same groups over time.* 

### Line charts by geography
#### High / low vulnerability
```{r, include=FALSE}
facet_line_by_geography_fixed <- function(geography, focus.text){
  # filter data
  data_clean_geo <- data_clean_more %>% 
    filter(county==geography)
  
  # make chart, adjust estimate type, subtitle
  static_facet_line_chart(t=data_clean_geo,
                          x="data_year_yr", y="fact_value",
                          fill="vulnerability", facet="focus_type_ord",
                          color="psrc_light",
                          est ="percent",
                          dec=2,
                          scales="fixed",
                          ncol=3,
                          title=paste0(indicator_title, focus.text),
                          subtitle="More than 1.5 persons per bedroom",
                          source="U.S. Census Bureau, American Community Survey (ACS) 5-Year Public Use Microdata Sample (PUMS)")+
    scale_y_continuous(limits = c(0, 30))+
    scale_y_continuous(labels = function(x) paste0(x*100, "%"))
}

facet_line_by_geography_free <- function(geography, focus.text){
  # filter data
  data_clean_geo <- data_clean_more %>% 
    filter(county==geography)
  
  # make chart, adjust estimate type, subtitle
  static_facet_line_chart(t=data_clean_geo,
                          x="data_year_yr", y="fact_value",
                          fill="vulnerability", facet="focus_type_ord",
                          color="psrc_light",
                          est ="percent",
                          dec=2,
                          scales="free",
                          ncol=3,
                          title=paste0(indicator_title, focus.text),
                          subtitle="More than 1.5 persons per bedroom",
                          source="U.S. Census Bureau, American Community Survey (ACS) 5-Year Public Use Microdata Sample (PUMS)")+
    scale_y_continuous(limits = c(0, 30))+
    scale_y_continuous(labels = function(x) paste0(x*100, "%"))
}
```

```{r}
facet_line_by_geography_fixed("Region", ": Region")
facet_line_by_geography_fixed("King", ": King County")
facet_line_by_geography_fixed("Kitsap", ": Kitsap County")
facet_line_by_geography_fixed("Pierce", ": Pierce County")
facet_line_by_geography_fixed("Snohomish", ": Snohomish County")

facet_line_by_geography_free("Region", ": Region")
facet_line_by_geography_free("King", ": King County")
facet_line_by_geography_free("Kitsap", ": Kitsap County")
facet_line_by_geography_free("Pierce", ": Pierce County")
facet_line_by_geography_free("Snohomish", ": Snohomish County")
```
\
\

#### calculated difference b/t
The 5 geographies are all included in the facets by geography, but they could be separated out to create 5 individual charts - one for each geography.
```{r}
dif_groups <- function(geography, group, category){
  region_btgrp_dif <- data_clean_more %>% 
    # dplyr::filter(county==geography) %>% 
    dplyr::filter(focus_type==group) %>% 
    group_by(data_year) %>% 
    mutate(dif = ifelse(focus_attribute!=category, fact_value - fact_value[focus_attribute == category], 0)) %>% 
    dplyr::filter(dif!=0)
}

poc <- dif_groups("Region", "POC_cat", 'POC')
dis <- dif_groups("Region", "Disability_cat", 'With disability')
income <- dif_groups("Region", "Income_cat", 'Low Income')
youth <- dif_groups("Region", "Youth_cat", 'Household with youth')
older <- dif_groups("Region", "Older_cat", 'Household with older adult')
lep <- dif_groups("Region", "LEP_cat", 'Limited English proficiency')

dif_equity_groups <- rbind(poc, 
                           dis, 
                           income,
                           youth,
                           older,
                           lep) %>% 
  dplyr::group_by(focus_type) # without this, colors were off

# facet line chart option
dif_equity <- static_facet_line_chart(t=dif_equity_groups,
                                      x="data_year_yr", y="dif",
                                      fill="county_ord", facet="focus_type_ord",
                                      color="psrc_light",
                                      est ="percent",
                                      dec=2,
                                      scales="fixed",
                                      ncol=3,
                                      title=paste0(indicator_title),
                                      subtitle="More than 1.5 persons",
                          source="U.S. Census Bureau, American Community Survey (ACS) 5-Year Public Use Microdata Sample (PUMS)") #+ theme(legend.position = "none")

dif_equity
```

### Line charts by equity group
#### High / low vulnerability
```{r, include=FALSE}
equity_group_facet_line <- function(focus.type, focus.text){
  # filter data
  data_clean_filter <- data_clean_more %>% 
    filter(focus_type==focus.type)
  
  # make chart, adjust estimate type, subtitle, and y limits
  static_facet_line_chart(t=data_clean_filter,
                          x="data_year_yr", y="fact_value",
                          fill="focus_attribute", facet="county",
                          color="psrc_light",
                          est ="percent",
                          dec=2,
                          scales="fixed",
                          ncol=3,
                          title=paste0(indicator_title, focus.text),
                          subtitle="More than 1.5 persons",
                          source="U.S. Census Bureau, American Community Survey (ACS) 5-Year Public Use Microdata Sample (PUMS)")+
    scale_y_continuous(limits = c(0, 30))+
    scale_y_continuous(labels = function(x) paste0(x*100, "%")) 
}
```

```{r}
equity_group_facet_line("POC_cat", " (race/ethnicity)")
equity_group_facet_line("Income_cat", " (income)")
equity_group_facet_line("Disability_cat", " (disability)")
equity_group_facet_line("Youth_cat", " (youth <18)")
equity_group_facet_line("Older_cat", " (adults 65+)")
equity_group_facet_line("LEP_cat", " (limited English proficiency)")
```
\
\

#### calculated difference b/t
The 6 equity focus groups are all included in the facets by geography, but they could be separated out to create 6 individual charts - one for each focus group.
```{r}
dif_groups <- function(group, category){

  btgrp_dif <- data_clean_more %>% 
    dplyr::filter(focus_type==group) %>% 
    group_by(data_year) %>% 
    mutate(dif = ifelse(focus_attribute!=category, fact_value - fact_value[focus_attribute == category], 0)) %>% 
    dplyr::filter(dif!=0)
}

poc <- dif_groups("POC_cat", 'POC')
dis <- dif_groups("Disability_cat", 'With disability')
income <- dif_groups("Income_cat", 'Low Income')
youth <- dif_groups("Youth_cat", 'Household with youth')
older <- dif_groups("Older_cat", 'Household with older adult')
lep <- dif_groups("LEP_cat", 'Limited English proficiency')

dif_equity_groups <- rbind(poc, 
                           dis, 
                           income,
                           youth,
                           older,
                           lep) %>% 
  dplyr::group_by(focus_type) # without this, colors were off

# facet line chart option
dif_equity <- static_facet_line_chart(t=dif_equity_groups,
                                      x="data_year_yr", y="dif",
                                      fill="focus_type_edit", facet="county_ord",
                                      color="psrc_light",
                                      est ="percent",
                                      dec=2,
                                      scales="fixed",
                                      ncol=3,
                                      title=paste0(indicator_title),
                                      subtitle="More than 1.5 persons",
                                      source="U.S. Census Bureau, American Community Survey (ACS) 5-Year Public Use Microdata Sample (PUMS)") #+ theme(legend.position = "none")

dif_equity
```
\

### Cleveland dot plot
[*Resource for visual*](https://uc-r.github.io/cleveland-dot-plots) 
\
The code to make this is type of visual is long - adjust to indicator as needed (*scale_x_continuous*, *labs*, etc).
\
```{r, edu-attainment, fig.width=10, eval = FALSE}
# get first and most recent years
data_clean_df <- data_clean_more %>% 
  dplyr::filter(data_year_yr==2011 |
                  data_year_yr==2021) %>% 
  dplyr::filter(county=="Region")

# order of y-axis
data_clean_df$focus_attribute_ord <- factor(data_clean_df$focus_attribute,
                                          levels = c("POC",
                                                     "Non-POC",
                                                     "Low Income",
                                                     "Non-Low Income",
                                                     "With disability",
                                                     "Without disability",
                                                     "Household with youth",
                                                     "Household without youth",
                                                     "Household with older adult",
                                                     "Household without older adult",
                                                     "Limited English proficiency",
                                                     "English proficient"))

# generate labels for data points - doesn't need to be adjusted
right_label <- data_clean_df %>%
  distinct(indicator_fact_id, .keep_all = TRUE) %>% 
  group_by(focus_attribute_ord) %>%
  arrange(desc(fact_value)) %>% 
  slice(1)

left_label <- data_clean_df %>%
  distinct(indicator_fact_id, .keep_all = TRUE) %>% 
  group_by(focus_attribute_ord) %>%
  arrange(desc(fact_value)) %>%
  slice(2)

# filter vulnerable groups of interest
highlight <- data_clean_df %>% 
  dplyr::filter(focus_attribute_ord=="Low Income" |
                  focus_attribute_ord=="Limited English proficiency" |
                  focus_attribute_ord=="With disability" |
                  focus_attribute_ord=="Household with youth" |
                  focus_attribute_ord=="Household with older adult" |
                  focus_attribute_ord=="POC")

# plot
p <- ggplot(data_clean_df, aes(fact_value, focus_attribute_ord)) +
  geom_line(aes(group = focus_attribute_ord), alpha=0.3) +
  geom_point(aes(color = data_year_yr, shape=data_year_yr), linewidth = 1.5, alpha=0.3)+
  scale_shape_manual(values = c(15, 16)) +
  scale_color_manual(values=c("#00A7A0","#F05A28")) +
  geom_line(data=highlight, aes(group = focus_attribute_ord), linewidth=0.9) +
  geom_point(data=highlight, aes(color = data_year_yr, shape=data_year_yr), linewidth = 2.1) +
  scale_shape_manual(values = c(15, 16)) +
  geom_text(data = right_label, aes(label = percent(fact_value, #adjust: percent or dollar
                                                    accuracy = 1)), #adjust to round
            linewidth = 4, hjust = -.5) +
  geom_text(data = left_label, aes(label = percent(fact_value, #adjust: percent or dollar
                                                   accuracy = 1)), #adjust to round
            linewidth = 4, hjust = 1.5) +
  scale_x_continuous(labels = scales::percent, #adjust: percent or dollar type
                     limits = c(0, 0.75)) + #these values depend on the indicator
  scale_y_discrete(limits=rev,
                   labels = function(focus_attribute_ord) stringr::str_wrap(focus_attribute_ord, width = 18))+
  theme_minimal() +
  theme(text=element_text(family="Poppins")) +
  labs(x="Percentage of adults", #following labs will need to be adjusted for indicator
       y=NULL,
       colour= "Year",
       shape="Year",
       title = paste0(indicator_title, ": Region"),
       subtitle = "More than 1.5 persons per bedroom",
       caption=paste("U.S. Census Bureau, American Community Survey (ACS) 5-Year Public Use Microdata Sample (PUMS)", "2011 disability data not available", sep = "\n")) +
  theme(plot.title = element_text(face="bold", linewidth=16),
        axis.text.y = element_text(face = c('plain','bold', 'plain', 'bold',
                                            'plain','bold', 'plain', 'bold')))
        

p
```
\
\

## 4. Save files
*This section needs to be edited. Keep the code chunks commented out for now as we draft and refine the visuals.*

### PNG
```{r}
# folder <- "Y:/Equity Indicators/experimenting/R-scripts/PUMS/output-visuals/"
# chart <- "kg-readiness.png"
# 
# ggsave(filename=paste0(folder,chart),
#        plot=p,
#        device='png',
#        width=5, height = 3, dpi = 300)
```

### HTML
```{r}
# interactive webpage output process
# htmlwidgets::saveWidget(healthrisk_quintile,
#                         file=paste0(website_dev_folder, data_nm,'.html'))
```

### Copy files from Y drive > website folder
```{r, include=FALSE, eval=FALSE}
# current.folder <- "Y:/Equity Indicators/tracker-webpage-content/economy"
# new.folder <- "//WEB/website_data/equity-tracker-webpages/economy"
# 
# list.of.files <- list.files(current.folder, full.names = T)
# list.of.files
# 
# # copy the files to the new folder
# file.copy(list.of.files,
#           new.folder,
#           overwrite = T)
```
\
\

## 5. Archive
*This section includes visuals that were determined to be less useful. We didn't want to lose the work, but didn't want to include it in the main workflow. Feel free to comment out if you don't want to adjust the arguments to fit the indicator of interest.*

### Line chart: all categories
```{r}
data_clean_region <- data_clean_more %>% 
  dplyr::filter(county=="Region")

#line_chart <- static_line_chart(t=data_clean_region,
 #                               x="data_year_yr", y="fact_value",
  #                              fill="focus_attribute_ord",
   #                             est = "percent",
    #                            breaks = NULL,
     #                           lwidth = 1,
      #                          color = "psrc_pairs",
       #                         title = "Housing cost burden: Region",
        #                        subtitle = "More than 1.5 persons per bedroom",
         #                       source = "U.S. Census Bureau, American Community Survey (ACS) 5-Year Public Use Microdata Sample (PUMS)") + ylim(0, 0.65)

#line_chart

line_chart <- interactive_line_chart(t=data_clean_region,
                                     x="data_year_yr", y="fact_value",
                                     fill="focus_attribute_ord",
                                     est = "percent",
                                     breaks = NULL,
                                     lwidth = 1,
                                     color = "psrc_pairs",
                                     title = "Housing cost burden: Region",
                                     subtitle = "More than 1.5 persons per bedroom",
                                     source = "U.S. Census Bureau, American Community Survey (ACS) 5-Year Public Use Microdata Sample (PUMS)")

line_chart
```
### Line chart: by vulnerability
```{r}
static_facet_line_chart(t=data_clean_region,
                        x="data_year_yr", y="fact_value",
                        fill="focus_type_ord", facet="vulnerability",
                        color="psrc_pairs",
                        est ="percent",
                        dec=2,
                        scales="fixed",
                        ncol=3,
                        title=paste0(indicator_title,": Region"),
                        subtitle="More than 1.5 persons per bedroom",
                        source="U.S. Census Bureau, American Community Survey (ACS) 5-Year Public Use Microdata Sample (PUMS)")
```
\
\

<a href="#top">Back to top of the page</a>