---
title: "census-permits-msa"
author: "suzanne"
date: "2023-08-11"
output:
  pdf_document: default
  html_document: default
---

In this task we need to download Excel-based and text based monthly Census reports of residential permit activity by MSA, and assembling the data for our two MSAs (Seattle-Tacoma-Bellevue and Bremerton-Silverdale-Port Orchard) into a consolidated monthly time-series dataset. 

Prior to Nov-2019, the files are text based, then switch to excel.

```{r}
library(readxl)
library(httr)
library(dplyr)
library(psrcplot)
library(ggplot2)
library(openxlsx)
library(stringr)
install_psrc_fonts()

```

## Get the data from January 2019 to November 2019 from text files

A bunch of specific unchanging parts of the file and url:
```{r}
col_names<-c("CSA", "CBSA", "Name", "Total", "1 Unit", "2 Units",
           "3 and 4 Units", "5 Units or More", 
           "Num of Structures With 5 Units or More", "Monthly Coverage Percent*", "Month")

url_start<-'https://www.census.gov/construction/bps/txt/tb3u2019'
url_end<-'.txt'
month_ids<-c('01', '02', '03', '04', '05', '06', '07', '08', '09', '10')
num_cols<-c("CSA", "CBSA","Total", "1 Unit", "2 Units",
           "3 and 4 Units", "5 Units or More", 
           "Num of Structures With 5 Units or More", "Monthly Coverage Percent*")
```

Get the data and clean it up
```{r}
download_clean_text<-function(url, month_id,colnames){
    big_tbl<-read.fwf(url, skip=11, header=FALSE, widths=c(3,6,40,8,8,8,8,8,8,8))
    
    # Select the region's MSA; the Bremerton one is 
    #spread across two lines and needs to be cleaned up
    psrc_msa_tbl<-big_tbl[c(47, 48,321),]
    
    col_names_no_month<-head(col_names, -1)
    
    
    psrc_msa_tbl<-setNames(psrc_msa_tbl, col_names_no_month)
    #clean up the Bremerton MSA data
    psrc_msa_tbl[1, 4:10]=psrc_msa_tbl[2,4:10]
    psrc_msa_tbl_clean<-psrc_msa_tbl[-2,]
    psrc_msa_tbl_clean$Month=paste0(2019,month_id)
return(psrc_msa_tbl_clean) 
}

```


put the months together
```{r}
df_2019<-data.frame(matrix(nrow = 0, ncol = length(col_names)))
colnames(df_2019)<-col_names

for(month_id in month_ids){
  url<-paste0(url_start, month_id, url_end)
  the_month_df<-download_clean_text(url, month_id,colnames)
  df_2019<-rbind(df_2019,the_month_df)
}
```

```{r}

# for being consistent with later datasets
df_2019<-df_2019%>%mutate_at(num_cols,as.numeric)%>%
  select(-c("Monthly Coverage Percent*"))
```

#Part two: Get data from November 2019 to June 2023


```{r}
month_ids<-c('01', '02', '03', '04', '05', '06', '07', '08', '09', '10', '11', '12')
yr_ids<-c('2020', '2021', '2022')

months_2019<-c('11','12')
months_2023<-c('01', '02', '03', '04', '05', '06')

month_yr_ids<-as.list(outer(yr_ids, month_ids, paste0))
months_2019_ids<-as.list(outer('2019', months_2019, paste0))
months_2023_ids<-as.list(outer('2023', months_2023, paste0))
month_yr_ids<-c(months_2019_ids,month_yr_ids,months_2023_ids)
url_start<-'https://www.census.gov/construction/bps/xls/msamonthly_'
url_end<-'.xls'

```

```{r}
download_clean_excel<-function(url, month_id, months_19_ids){

    GET(url, write_disk(tf<-tempfile(fileext=".xls")))
  # sheet names changed after 2019
    if(month_id %in% months_2019_ids){
      sheetName='Units'}
    else{
      sheetName='MSA Units'
    }
  
    big_tbl<-read_excel(tf, sheet=sheetName,range='A8:I377')
    
    psrc_msa_tbl_clean<-big_tbl%>%filter(Name %in% c('Seattle-Tacoma-Bellevue, WA', 'Bremerton-Silverdale-Port Orchard, WA'  ))
    psrc_msa_tbl_clean$Month=month_id
return(psrc_msa_tbl_clean) 
}
```





```{r}
col_names<-col_names[names(col_names)!="Monthly Coverage Percent*"]
df_new<-data.frame(matrix(nrow = 0, ncol = length(col_names)))
colnames(df_new)<-col_names


for(month_id in month_yr_ids){
  url<-paste0(url_start, month_id, url_end)
  print(url)
  the_month_df<-download_clean_excel(url, month_id, months_19_ids)
  df_new<-rbind(df_new,the_month_df)
}
```


```{r}
df_all<-rbind(df_2019, df_new)
```

```{r}
df_all<-df_all%>%mutate(MSA_Name=str_trim(Name))

```

```{r}
write.xlsx(df_all, 'census_bps_monthly.xlsx')
```

```{r}
static_line_chart(t=df_all, x='Month', y='Total', fill='MSA_Name',title='Total Permits by Month Nov 2019-June 2023' )+ theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))+ theme(legend.position = "bottom")
```
```{r}
static_line_chart(t=df_all, x='Month', y='1 Unit', fill='MSA_Name',  title='1 Unit Permits by Month Nov 2019-June 2023')+ theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))+ theme(legend.position = "bottom")
```

```{r}
static_line_chart(t=df_all, x='Month', y='2 Units', fill='MSA_Name',title='2 Unit Permits by Month Nov 2019-June 2023')+ theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))+ theme(legend.position = "bottom")
```

```{r}
static_line_chart(t=df_all, x='Month', y='3 and 4 Units', fill='MSA_Name', title='3 and 4 Unit Permits by Month Nov 2019-June 2023')+ theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))+ theme(legend.position = "bottom")
```

```{r}
static_line_chart(t=df_all, x='Month', y='5 Units or More', fill='MSA_Name', title='5 or More Units permits by month Nov 2019-June 2023')+ theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))+ theme(legend.position = "bottom")
```

```{r}
static_line_chart(t=df_all, x='Month', y='Num of Structures With 5 Units or More', title='Num of Structures With 5 Units or More by Month', fill='MSA_Name')+ theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))+ theme(legend.position = "bottom")
```
